\section{Multimodal features in fashion recommendation}

\subsection{Visual Features}

In fashion recommendation, visual attributes (such as color, pattern, style, shape, and fabric texture) are critical determinants of user preferences. Because clothing decisions are heavily guided by aesthetics, purely collaborative signals or textual descriptions often fail to capture these stylistic nuances. Consequently, incorporating visual features is essential to enrich item representations \parencite{10.1007/978-981-95-3355-8_6}. In this study, visual features are processed and stored as \texttt{image\_embeddings.npy} using two alternative encoders:
\begin{itemize}
    \item \textit{MobileNetV2}: A lightweight convolutional neural network pre-trained on ImageNet. It processes the item image \(I_i\) to extract a spatial feature representation:
    \begin{equation}
        \mathbf{x}_i^{\text{vis}} = \text{MobileNetV2}(I_i) \in \mathbb{R}^{d_v},
    \end{equation}
    where \(d_v\) is the visual feature dimension. MobileNetV2 is highly computationally efficient, making it suitable for low-latency inference and resource-constrained deployments, although it lacks deep semantic alignment with textual concepts.
    \item \textit{CLIP (Contrastive Language-Image Pretraining)}: A state-of-the-art vision-language model trained on web-scale image-text pairs. The visual encoder of CLIP maps the image \(I_i\) into a joint vision-language latent space:
    \begin{equation}
        \mathbf{x}_i^{\text{vis}} = \text{CLIP}_{\text{vis}}(I_i) \in \mathbb{R}^{d_v}.
    \end{equation}
    While computationally heavier than MobileNetV2, CLIP features capture high-level semantic abstractions, representing aesthetic style and textual compatibility in a unified space.
\end{itemize}
The trade-off between these encoders is clear: CLIP provides richer semantic representations at the cost of higher feature extraction overhead, whereas MobileNetV2 offers a lightweight and fast alternative for standard convolutional feature extraction.


\subsection{Textual Features}

Textual metadata, such as product titles, descriptions, categories, and tags, provides explicit, descriptive information about fashion items. In this study, the textual modality is represented in two complementary ways depending on the integration paradigm:
\begin{itemize}
    \item \textit{BERT (Bidirectional Encoder Representations from Transformers)}: For the model-level architectures (BM3 and FREEDOM), item text is encoded by a pre-trained \texttt{bert-base-uncased} model into a dense semantic embedding \(\mathbf{x}_i^{\text{txt}} \in \mathbb{R}^{768}\), stored as \texttt{text\_embeddings.npy}. These contextual embeddings capture semantic relatedness beyond exact keyword overlap (e.g., associating ``dress'' with ``gown'').
    \item \textit{TF-IDF (Term Frequency-Inverse Document Frequency)}: For the data-level architecture (CombiGCN), item text is instead represented by sparse lexical vectors \(\in \mathbb{R}^{V}\) (where \(V\) is the vocabulary size), used only to construct the item-item textual similarity graph via cosine similarity.
\end{itemize}
Thus BERT provides the dense textual feature consumed directly by the model-level encoders, while TF-IDF provides the lexical similarity prior used by CombiGCN's similarity graph.


\subsection{Multimodal Fusion}

To leverage both visual aesthetics and textual attributes, multimodal recommender systems employ fusion strategies to integrate features from different modalities into a unified item representation. In the proposed pipeline, the choice of fusion is parameterized by the \texttt{sim\_type} variable, evaluating the following configurations:
\begin{itemize}
    \item \textit{Visual-only (\texttt{img\_only})}: The item representation is constructed solely from the visual embedding: \(\mathbf{x}_i = \mathbf{x}_i^{\text{vis}}\).
    \item \textit{Textual-only (\texttt{text\_only})}: The item representation is constructed solely from the textual embedding: \(\mathbf{x}_i = \mathbf{x}_i^{\text{txt}}\). This setting corresponds to the \texttt{tfidf} flag in the implementation; it is realized as the BERT text embedding for BM3/FREEDOM and as TF-IDF lexical similarity for CombiGCN.
    \item \textit{Late Fusion (\texttt{multimodal})}: A parameter-free fusion approach that serves as a stable baseline. It directly combines the visual and textual representations (e.g., via vector concatenation or element-wise averaging) without requiring additional learnable parameters:
    \begin{equation}
        \mathbf{x}_i = \mathbf{x}_i^{\text{vis}} \oplus \mathbf{x}_i^{\text{txt}},
    \end{equation}
    where \(\oplus\) denotes the concatenation operator.
    \item \textit{Weighted Attention Fusion (\texttt{multimodal\_attention})}: An adaptive fusion mechanism that dynamically learns attention weights to balance the contributions of visual and textual modalities on a per-item basis. This approach utilizes trainable projection layers and an attention gating network to automatically prioritize the most descriptive modality. As analyzed in Chapter 4, however, the additional trainable parameters introduced by this mechanism do not consistently improve recommendation accuracy under sparse interaction data:
    \begin{equation}
        \mathbf{x}_i = \beta_{v,i} W_v \mathbf{x}_i^{\text{vis}} + \beta_{t,i} W_t \mathbf{x}_i^{\text{txt}},
    \end{equation}
    where \(W_v \in \mathbb{R}^{d_m \times d_v}\) and \(W_t \in \mathbb{R}^{d_m \times d_t}\) are trainable projection matrices mapping visual and textual features to a shared \(d_m\)-dimensional space, and \(\beta_{v,i}\) and \(\beta_{t,i}\) are visual and textual attention coefficients satisfying \(\beta_{v,i} + \beta_{t,i} = 1\).
\end{itemize}

Advanced multimodal recommendation models integrate these fused features directly into their learning objectives:
\begin{itemize}
    \item \textit{BM3 \parencite{10.1145/3543507.3583251}}: Uses modality-specific projection layers and optimizes a self-supervised bootstrap contrastive loss to align user and item multimodal representations without requiring negative samples.
    \item \textit{FREEDOM}: Constructs a frozen item-item kNN similarity graph using item multimodal features, and propagates content embeddings over this static structure, alleviating the computational burden of dynamic graph convolutions on user-item bipartite graphs.
\end{itemize}

Despite the variety of multimodal fusion and graph-based models, there remains a lack of systematic evaluation comparing different fusion configurations (\texttt{img\_only}, \texttt{text\_only}, \texttt{multimodal}, \texttt{multimodal\_attention}) across graph neural collaborative filtering architectures built on the LightGCN scheme (CombiGCN, BM3, and FREEDOM) in the fashion domain. This study addresses this gap by conducting a comprehensive comparative analysis of these modalities and fusion mechanisms under diverse structural propagation environments.
