\section{Conclusion}

This study has investigated the integration of heterogeneous multimodal features with graph-based collaborative filtering architectures to solve the challenges of extreme interaction sparsity in personalized fashion recommendation. Through a systematic evaluation of 24 model configurations—spanning three graph neural network (GNN) architectures (BM3, CombiGCN adapted, and FREEDOM), two vision encoders (MobileNetV2 and CLIP), a BERT textual encoder, and four similarity and fusion strategies—several key empirical and theoretical conclusions can be drawn:

First, within every evaluated multimodal graph architecture, fusing visual and textual features via late fusion consistently outperforms its single-modality variants (visual-only and textual-only). By propagating aesthetic compatibility and textual semantics along item-item similarity links, multimodal graph networks strengthen the representations of low-interaction items relative to their unimodal counterparts.

Second, the visual feature extractor plays a critical role in style compatibility representation. In the best-performing configuration of each architecture, MobileNetV2 (MBNv2) matches or outperforms the CLIP vision-language foundation model. We emphasize that this superiority is configuration-dependent rather than universal: under the visual-only (\texttt{img\_only}) setting of CombiGCN, CLIP substantially outperforms MobileNetV2 (NDCG@20 of $0.0220$ versus $0.0119$). The advantage of MobileNetV2 therefore holds specifically for the late-fusion regime on this dataset. A plausible explanation is that while CLIP is optimized for high-level semantic cross-modal alignment (e.g., classifying broad categories like a ``red jacket''), fashion styling compatibility is also governed by fine-grained visual details such as local textures, stitching, fabric patterns, and collar lines. MobileNetV2, utilizing local spatial convolutions, preserves these low-level spatial styling features, making it well suited to the late-fusion regime for visual garment compatibility.

Third, the selection of a multimodal fusion strategy must be balanced against dataset scale. Late fusion (\texttt{multimodal}) via element-wise average pooling represents a robust, parameter-efficient baseline that consistently yields peak performance. Conversely, trainable weighted attention gating (\texttt{multimodal\_attention}) introduces excessive model complexity and overfitting on sparse, small-scale datasets (such as VCR), degrading the performance of robust models like BM3 and CombiGCN while only slightly assisting weaker baselines (FREEDOM).

Finally, model performance is highly retrieval-depth dependent. BM3, utilizing self-supervised bootstrap representation alignment, is the best-performing framework among those evaluated for list sizes $K \ge 5$. The adapted CombiGCN dual-graph model remains competitive at the shallowest retrieval depths---tying BM3 at top-1 (NDCG@1 $=$ Precision@1 $= 0.0127$ for both) and marginally leading on top-10 precision---but BM3 leads on all ranking-quality metrics from $K \ge 5$. These findings demonstrate that visual-textual GNNs are highly effective for personalized fashion recommendation, and their structural components must be carefully optimized for the target application scenario.
