\chapter*{INTEGRATING MULTIMODAL REPRESENTATIONS INTO GRAPH-BASED FASHION RECOMMENDER SYSTEMS\\ABSTRACT}

Personalized fashion recommender systems face severe challenges due to high user-item interaction sparsity and the limitations of ID-bound collaborative filtering models. To address these bottlenecks, this study investigates multimodal graph collaborative filtering by integrating visual and textual representations of clothing items into Graph Neural Network (GNN) architectures. Building on LightGCN as the shared collaborative-filtering backbone, we systematically benchmark three multimodal graph architectures---CombiGCN (adapted to operate over item similarity graphs), BM3, and FREEDOM---under twenty-four model configurations. These configurations utilize diverse feature encoders (MobileNetV2, CLIP, and TF-IDF) and fusion paradigms (late fusion average pooling vs. weighted attention gating) on the Vibrent Clothes Rental (VCR) dataset. Empirical results reveal three key findings. First, within the best-performing configuration of each architecture, the MobileNetV2 vision encoder matches or outperforms the CLIP vision-language foundation model, suggesting that localized spatial texture descriptors are effective for garment styling compatibility; we note, however, that this advantage is configuration-dependent rather than universal, as CLIP surpasses MobileNetV2 under certain visual-only settings. Second, the parameter-free late fusion strategy serves as a highly stable and robust baseline; in contrast, introducing trainable parameters via weighted attention gating induces semantic noise and overfitting when applied to sparse interaction logs. Third, model superiority is retrieval-depth dependent: BM3, utilizing self-supervised contrastive bootstrap alignment, attains the best performance among the evaluated models for ranking lists of size $K \ge 5$, whereas the adapted CombiGCN dual-graph model remains competitive at the shallowest retrieval depths, tying BM3 at top-1 and marginally leading on top-10 precision. Overall, this study indicates that embedding multimodal features into GNNs helps mitigate data sparsity within the evaluated model family, providing empirical guidelines for designing personalized fashion recommendation systems.