\section{Experimental setup}

This section details the experimental configuration, hyperparameter selections, and evaluation protocols established to compare the recommender system models under uniform conditions.

\subsection{Configuration Space}

To systematically evaluate the impact of different visual/textual encoders and modal fusion strategies across different graph structures, we define a grid of 24 model configurations. Specifically, the configuration space is spanned by three dimensions:
\begin{itemize}
    \item \textbf{Recommendation Models}: We evaluate three graph collaborative filtering architectures: CombiGCN (adapted) \parencite{10.1007/978-981-97-0669-3_11}, BM3 \parencite{10.1145/3543507.3583251}, and FREEDOM \parencite{Zhou_2023}. All three build on the LightGCN \parencite{he2020lightgcnsimplifyingpoweringgraph} propagation scheme as their shared collaborative-filtering backbone; LightGCN itself is not run as a standalone configuration.
    \item \textbf{Feature Encoders}: We evaluate two vision encoders (MobileNetV2 vs. CLIP) for the visual modality. The textual modality is encoded by BERT (\texttt{bert-base-uncased}, $768$-d) as dense embeddings for the model-level architectures (BM3, FREEDOM), and by TF-IDF lexical similarity for the data-level CombiGCN graph. Raw features are stored as \texttt{image\_embeddings.npy} and \texttt{text\_embeddings.npy} respectively.
    \item \textbf{Fusion and Similarity Types (\texttt{sim\_type})}: We evaluate four settings:
    \begin{itemize}
        \item \texttt{img\_only}: Item representation uses visual embeddings only.
        \item \texttt{text\_only}: Item representation uses textual embeddings only (BERT for BM3/FREEDOM; TF-IDF similarity for CombiGCN). This setting is labelled \texttt{tfidf} in the implementation.
        \item \texttt{multimodal}: Late fusion combining visual and textual features via average pooling or concatenation.
        \item \texttt{multimodal\_attention}: Trainable weighted attention fusion dynamically gating visual and textual contributions.
    \end{itemize}
\end{itemize}

Note that the \texttt{text\_only} setting is textual-only and therefore independent of the choice of vision encoder; for presentation symmetry it is reported under both the CLIP and MobileNetV2 encoder groups, so the $24$ reported cells correspond to $18$ distinct model--feature combinations.

\subsection{Research Questions}

The experimental analysis in Section 4.3 is organized around four research questions (RQs), each isolating one design factor (encoder, modal fusion strategy, and architecture) to determine which factor governs recommendation quality:
\begin{itemize}
    \item \textbf{RQ1 (Encoder)}: Which visual encoder, CLIP or MobileNetV2, is more suitable for the fashion recommendation domain?
    \item \textbf{RQ2 (Fusion strategy)}: Which feature integration strategy is most effective? In particular, does trainable weighted attention gating (\texttt{multimodal\_attention}) actually improve upon parameter-free late fusion (\texttt{multimodal})?
    \item \textbf{RQ3 (Model)}: Among the three architectures (BM3, CombiGCN, and FREEDOM), which performs best when each is evaluated under its own optimal configuration?
    \item \textbf{RQ4 (Consistency)}: Is the resulting performance ranking stable across different cut-off depths $K$ and across evaluation metrics?
\end{itemize}

\subsection{Key Hyperparameters}

To ensure reproducibility, all models are trained using uniform optimization hyperparameters where applicable. The common parameters are defined as follows:
\begin{itemize}
    \item Embedding dimension ($d$): 512 for user/item ID embeddings and projected features.
    \item Number of GNN layers ($K$): 4 layers of graph propagation.
    \item Learning rate ($\eta$): $0.001$, optimized via the Adam optimizer.
    \item L2 regularization decay coefficient ($\lambda$): $10^{-4}$.
    \item Batch size: 8,192 interactions per mini-batch.
    \item Epochs and stopping criteria: All models are trained for a maximum of 1,000 epochs. Early stopping is applied with a patience of 40 epochs, monitored via Recall@20 on the test set.
\end{itemize}
Model-specific contrastive learning parameters are configured as:
\begin{itemize}
    \item \textbf{BM3}: Contrastive learning weight $\lambda_{CL} = 0.2$, target EMA encoder momentum $m = 0.995$.
    \item \textbf{FREEDOM}: Contrastive learning weight $\lambda_{CL} = 0.1$, InfoNCE temperature $\tau = 0.2$, and kNN graph size $k = 10$.
\end{itemize}

\subsection{Choice of K Values and Rationale}

To measure top-$N$ recommendation accuracy, candidate items for each user are ranked in descending order of their predicted scores. We evaluate recommendation performance at four different cut-off values: $K \in \{1, 5, 10, 20\}$. 

The choice of $K=5$ is designated as the primary evaluation setting in our analysis. The technical rationale is based on the characteristics of the VCR test set: the test set contains an average of only $3.81$ items per user (see Table~\ref{tab:dataset_stats}). Consequently, a recommendation list of size $K=10$ or $K=20$ is larger than the user's ground-truth catalog, which leads to recall saturation and compresses the performance margins between different configurations. Evaluating at $K=5$ provides a tighter target window (approximately $1.3\times$ the size of the ground truth), which effectively prevents metric saturation and allows for a clearer differentiation of model performance. The values $K=10$ and $K=20$ are maintained to facilitate comparisons with standard GNN literature.

We report performance across six standard top-$N$ recommendation metrics: Recall@K, Normalized Discounted Cumulative Gain (NDCG@K), and Hit Ratio (HR@K) as our primary metrics, while Precision@K, Mean Average Precision (MAP@K), and Mean Reciprocal Rank (MRR@K) serve as secondary metrics.
